---
title: "Sleep duration regression CLEAR-3"
output:
  html_document:
    toc: yes
    df_print: paged
  pdf_document:
    toc: yes
  html_notebook:
    toc: yes
    df_print: paged
    toc_float: yes
    theme: spacelab
---

This R notebook takes the GRSVD values generated in the Jupyter notebook and uses them to predict Oura-derived sleep duration, among other things. The organisation is as follows:

1. Reading data
2. Generating plots and some descriptive statistics
3. Running and evaluating linear mixed-effects models to predict 1) self-reported sleep duration, 2) Oura-derived time-in-bed, and 3) Oura-derived sleep duration
4. Doing some miscellaneous work for e.g. sensitivity analyses

To be able to run this notebook, install the packages listed below using `install.packages("<package-name>")` (e.g., `install.packages("haven")`). Make sure that `utils.R`, which is available from the GitHub repository and contains some utility functions, is within the same folder as this file. Finally, create a .env text file that specifies the data and manuscript image directories, and make sure these directories exist:

```
DAT_DIR=/path/to/data/directory
MAN_IMG_DIR=/path/to/image/directory
```

Once all that is done, the setup chunk can be run:

```{r setup, echo=TRUE, results='hide', warning=FALSE, message=FALSE}
library(haven)
library(MuMIn)
library(arrow)
library(tidyverse)
library(lme4)
library(nlme)
library(ggeffects)
library(rlang)
library(tableone)
library(r2mlm)
library(ggpubr)

library(conflicted)
conflicts_prefer(dplyr::select, dplyr::filter, dplyr::lag, rlang::`%@%`)

source("utils.R")

# Read environment variables from a .env file
readRenviron(".env")

dat_dir <- Sys.getenv("DAT_DIR") # Data directory
man_img_dir <- Sys.getenv("MAN_IMG_DIR") # Manuscript image directory

# Otherwise hour and date extraction use local time
Sys.setenv(TZ='UTC')
```

# Reading data

This section loads demographics, Oura ring, and GRSVD data.

## Demographics

Demographics are split between a controls and intervention file.

```{r echo=TRUE, results='hide'}
dem_ctrl <- read_sav(file.path(dat_dir, "CLEAR3_Biaffect_Leow", "clear123ctrls_demotable.sav")) %>%
  as_factor() %>%
  # Don't need much more right now
  select(id, age, houseincomecat, gender) %>%
  filter(id >= 3500) %>%
  mutate(id = as.character(id)) %>%
  arrange(id) %>%
  add_column(control = TRUE)

dem_ctrl
```


```{r echo=TRUE, results='hide'}
dem <- readRDS(file.path(dat_dir, "CLEAR3_Biaffect_Leow", "clear3trait.rds")) %>%
  as_factor() %>% # Convert haven_labelled variables to factors
  mutate(
    id = as.character(id),
    race_clear123 = as_factor(race_clear123),
    ethnicity = as_factor(ethnicity)
  ) %>%
  arrange(id) %>%
  mutate(across(where(is.character), ~ if_else(.x == "", "Unknown", .x))) %>%
  add_column(control = FALSE) %>%
  bind_rows(dem_ctrl) %>%
  rename(Age = age, Gender = gender, `House income` = houseincomecat)

# Given to me by Anisha
dem[dem$id == "3085",]$Age <- 24
dem[dem$id == "3100",]$Age <- 23

dem
```

## Oura ring

```{r echo=TRUE, results='hide'}
oura_tib <- readRDS(file.path(
  dat_dir, "CLEAR3_Biaffect_Leow", "clear3baseline_extrasleepvars.rds")) %>%
  ungroup() %>%
  drop_na(TimeInBed_SRI) %>%
  mutate(
    TimeInBed_hour = TimeInBed_SRI / 3600,
    daterated = as_date(daterated),
    delta_start = Oura_bedtime_start_delta / 3600,
    delta_end = Oura_bedtime_end_delta / 3600
  )

dat_oura <- readRDS(
  file.path(dat_dir, "CLEAR3_Biaffect_Leow", "clear3baseline_sleep.rds")) %>%
  ungroup() %>%
  full_join(oura_tib, by = c("id", "daterated")) %>%
  mutate(
    Oura_Sleep_hour = Oura_Sleep_TST / 3600, # Second to hour
    sleepdur_yest = sleepdur_yest + 1 # Likert to hours
  ) %>%
  filter(id != 3086 | daterated != ymd("2022-07-31")) %>% # Duplicate entries
  mutate(TimeInBed_hour = case_when(
    TimeInBed_hour < 0 ~ NA_real_,
    TRUE ~ TimeInBed_hour
  )) # Exclude negative entries for now

dat_oura
```

The min and max number of hours recorded by Oura.

```{r}
range(dat_oura$Oura_Sleep_hour, na.rm = TRUE)
```


```{r eval=FALSE, include=FALSE}
tmp <- dat_oura %>%
  mutate(
    start_24 = delta_start + 24,
    end_24 = delta_end + 24,
    start_wrapped = case_when(
      start_24 > 24 ~ start_24 - 24,
      TRUE ~ start_24
    ),
    end_wrapped = case_when(
      end_24 > 24 ~ end_24 - 24,
      TRUE ~ end_24
    )
  )

hist(tmp$start_wrapped, breaks = 24)
hist(tmp$end_wrapped, breaks = 24)

# Note that this requires dat_oura to still have the negative TIBs

tmp2 <- tmp %>%
  filter(TimeInBed_hour < 0) %>%
  mutate(position = 4.5 + seq(0, 0.5, length.out = n())) %>%
  pivot_longer(c(start_wrapped, end_wrapped), names_to = "variable")

ggplot(tmp2, aes(value)) +
  geom_histogram(position = "dodge", color = "white") +
  geom_point(aes(color = as.factor(id), y = position)) +
  labs(x = "Time of day", y = "Count") +
  facet_wrap(
    ~ factor(variable, levels = c("start_wrapped", "end_wrapped")), 
    labeller = as_labeller(c(
      end_wrapped = "End time", 
      start_wrapped = "Start time"
    ))
  )
```


```{r}
cat("Number of participants in dat_oura:\n")
n_distinct(dat_oura$id)
cat("Number of participants with Oura measurements:\n")
n_distinct(drop_na(dat_oura, Oura_Sleep_TST)$id)
cat("Number of participants with self-report of sleep duration:\n")
n_distinct(drop_na(dat_oura, sleepdur_yest)$id)
```

Convert to parquet to make accessible to Python.

```{r eval=FALSE, include=TRUE}
dat_oura_raw <- readRDS(
  file.path(dat_dir, "CLEAR3_Biaffect_Leow", "clear3baseline_sleep.rds"))

write_parquet(dat_oura_raw, file.path(dat_dir, 
                                      "CLEAR3_Biaffect_Leow", 
                                      "clear3baseline_sleep.parquet"))
```


```{r fig.height=10}
ggplot(dat_oura, aes(sleepdur_yest, Oura_Sleep_hour)) +
  geom_jitter() +
  geom_abline(intercept = 0, slope = 1) +
  labs(title = "Recorded vs reported sleep time across participants (jittered)") +
  xlab("Self-reported sleep duration (hour)") +
  ylab("Oura-recorded total sleep time (hour)") +
  theme(text = element_text(size = 18))

ggplot(dat_oura, aes(SleepLNQuality, Oura_Sleep_hour)) +
  geom_point() +
  labs(title = "Recorded sleep time vs reported sleep quality across participants") +
  xlab("Self-reported sleep quality (higher is better)") +
  ylab("Oura-recorded total sleep time (hour)") +
  theme(text = element_text(size = 18))

ggplot(dat_oura, aes(sleepdur_yest, SleepLNQuality)) +
  geom_jitter() +
  labs(title = "Reported sleep time vs quality across participants") +
  xlab("Self-reported sleep duration (hour)") +
  ylab("Self-reported sleep quality (higher is better)") +
  theme(text = element_text(size = 18))
```

Correlation between self-report and Oura ring sleep duration.

```{r}
cor.test(dat_oura$sleepdur_yest, dat_oura$Oura_Sleep_hour, method = "spearman")
```

## GRSVD values

This code creates BiAffect-derived sleep durations from the GRSVD values. This is done by thresholding the values and counting the number of hours that fall below the threshold. Thresholds can either be data-driven (e.g., Otsu's method or a mixture of Gaussians) or fixed by the experimenter.

Note that this function also does some time shifting (+12 hours) to make sure the night ends up in the middle of every 24-hour period. This allows for easier counting of the number of hours slept per date.

```{r echo=TRUE, results='hide'}
sleep_preds_svd <- function(file, dat_dir, dat_oura, sleep_threshold = 1) {
  # Based on CLEAR3_processed_2024-10-24/dat_kp.parquet
  hourly_tz <- read_parquet(file.path(
    dat_dir,
    "CLEAR3_processed_2024-10-24", 
    "hourly_time_zone.parquet"
  ))
  
  read_parquet(file) %>%
    relocate(subject, .before = 1) %>%
    mutate(
      hour = hour(timestamp),
      date = date(timestamp)
    ) %>%
    left_join(hourly_tz, 
              by = c("subject", "timestamp" = "roundedTimestampUTC")) %>%
    group_by(subject) %>%
    fill(timezone, .direction = "downup") %>%
    group_by(subject, date) %>%
    mutate(
      n_timezones = n_distinct(timezone),
    ) %>%
    filter(n_timezones == 1) %>% # Exclude days with a time zone switch
    mutate(
      timestamp_local = timestamp + hours(timezone / 100),
      timestamp_warped = timestamp_local + hours(12), # Center the night
      hour_warped = hour(timestamp_warped),
      date_warped = date(timestamp_warped)
    ) %>%
    relocate(timestamp_warped, .after = timestamp) %>%
    mutate(sleep = svd_value < sleep_threshold) %>%
    group_by(subject, date_warped) %>%
    filter(n() == 24) %>% # Only allow complete rows in the warped matrix
    summarize(
      hour_estimate = sum(sleep), # Don't remove NAs
      hour_estimate_otsu = sum(svd_value < threshold_otsu),
      hour_estimate_gm = sum(svd_value < threshold_gm),
      activity = sum(svd_value),
      n_total_presses = sum(n_presses),
      n_active_hours = sum(n_presses > 0),
      label = first(label)
    ) %>%
    rename(date = date_warped) %>%
    ungroup() %>%
    full_join(mutate(dat_oura, id = as.character(id)), 
              by = c("subject" = "id", "date" = "daterated")) %>%
    filter(subject != "3086" | date != ymd("2022-07-31")) %>% # Duplicate entries
    ungroup()
}


dat <- sleep_preds_svd(
  file.path(
    dat_dir, 
    "svd_mats_long_alpha1_2025-07-30_utc-reference_min30.parquet"
  ),
  dat_dir,
  dat_oura
)

dat_dem <- dat %>%
  left_join(dem, by = c("subject" = "id"))

dat_cmplt_sr <- dat_dem %>%
  drop_na(hour_estimate, sleepdur_yest)

dat_cmplt_oura <- dat_dem %>%
  drop_na(c(hour_estimate, Oura_Sleep_hour))

dat
```

```{r}
n_distinct(dat_cmplt_sr$subject)
n_distinct(dat_cmplt_oura$subject)
```

The GRSVD parameters were fitted on a small portion of each participant's BiAffect data; this portion is called the training set. The rest of the data is called the test set and was collected after training set collection. We want to exclude the training set when fitting our mixed models. Due to autocorrelation, there might additionally be some training set information that leaked into the test set. For that reason, we also exclude the first few days of the test set (amounts to a total of 115).

```{r echo=TRUE, results='hide'}
# We can only reliably find the last training days in the full data set
last_train_days <- dat %>%
  drop_na(date) %>%
  filter(label == "train") %>%
  group_by(subject) %>%
  summarize(last_train_day = tail(date, n = 1))

# Helper function
filter_mutate <- function(.x) {
  .x %>%
    left_join(last_train_days, by = "subject") %>%
    filter(label == "test", date > last_train_day + days(3)) %>% # Don't include first three days in test region
    mutate(
      activity_c = scale(activity),
      hour_estimate_c = drop(scale(hour_estimate)),
      hour_estimate_otsu_c = drop(scale(hour_estimate_otsu)),
      hour_estimate_gm_c = drop(scale(hour_estimate_gm)),
      sleepdur_yest_c = drop(scale(sleepdur_yest, scale = FALSE)),
      age_c = scale(Age)
    ) %>%
    group_by(subject) %>%
    mutate(
      # Not scaling to unit variance to avoid throwing away information
      hour_estimate_cluster_centered = drop(scale(hour_estimate, scale = FALSE)),
      hour_estimate_cluster_mean = mean(hour_estimate, na.rm = TRUE)
    ) %>%
    ungroup() %>%
    mutate( # Scale everything again to help fitting
      hour_estimate_cluster_centered = drop(scale(hour_estimate_cluster_centered)),
      hour_estimate_cluster_mean = drop(scale(hour_estimate_cluster_mean))
    )
}

dat_test_sr <- dat_cmplt_sr %>%
  filter_mutate()

dat_test_oura <- dat_cmplt_oura %>%
  filter_mutate()

dat_test_oura
```

```{r}
n_distinct(dat_test_sr$subject)
n_distinct(dat_test_oura$subject)
```

# Descriptive statistics

Some plots to gain a feel for how many data we have per participant, what the data distributions are, and how some variables relate to each other.

```{r fig.height=8, fig.width=8}
ggplot(dat_oura %>% drop_na(Oura_Sleep_TST), aes(id)) +
  geom_bar() +
  labs(x = "Participant ID", y = "Count") +
  theme(
    text = element_text(size = 18),
    axis.text.x = element_text(angle = 90, vjust = 0.5)
  )

ggplot(dat_test_oura, aes(subject)) +
  geom_bar() +
  labs(x = "Participant ID", y = "Count") +
  theme(
    text = element_text(size = 18),
    axis.text.x = element_text(angle = 90, vjust = 0.5)
  )
```

Self-report.

```{r fig.height=8}
ggplot(dat_test_sr, aes(sleepdur_yest)) +
  geom_histogram() +
  labs(x = "Self-reported sleep duration")

ggplot(dat_test_sr, aes(SleepLNQuality)) +
  geom_histogram() +
  labs(x = "Self-reported sleep quality")

ggplot(dat_test_sr, aes(activity)) +
  geom_histogram() +
  labs(x = "Activity score")

ggplot(dat_test_sr, aes(hour_estimate)) +
  geom_histogram()

ggplot(dat_test_sr, aes(activity_c, sleepdur_yest)) +
  geom_jitter()

ggplot(dat_test_sr, aes(hour_estimate, sleepdur_yest)) +
  geom_jitter() +
  geom_abline(intercept = 0, slope = 1) +
  labs(title = "Reported vs estimated sleep (jittered)", 
       x = "Hour estimate", y = "Self-reported sleep duration") +
  theme(text = element_text(size=16))

ggplot(dat_test_sr, aes(hour_estimate, SleepLNQuality)) +
  geom_jitter(alpha = 0.2) +
  labs(title = "Reported quality vs estimated duration (jittered)", 
       x = "Hour estimate", y = "Self-reported sleep quality") +
  theme(text = element_text(size=16))
```

Oura ring:

```{r fig.height=8}
n_presses_breaks = c(400, 1100, 3000, 8100, 22000)
ggplot(dat_test_oura, aes(hour_estimate, Oura_Sleep_hour)) +
  geom_jitter(aes(color = n_total_presses)) +
  geom_abline(intercept = 0, slope = 1) +
  scale_color_gradient(name = "Number of key presses", trans = "log",
                       breaks = n_presses_breaks, labels = n_presses_breaks) +
  labs(title = "Recorded vs estimated sleep (jittered)", 
       x = "Hour estimate", y = "Oura sleep duration") +
  theme(text = element_text(size=16))

ggplot(dat_test_oura, aes(hour_estimate, Oura_Sleep_hour)) +
  geom_jitter(aes(color = n_active_hours)) +
  geom_abline(intercept = 0, slope = 1) +
  labs(title = "Recorded vs estimated sleep (jittered)", 
       x = "Hour estimate", y = "Oura sleep duration") +
  theme(text = element_text(size=16))
```

Self-report and Oura ring:

```{r}
# 1887 observations
tmp <- dat_oura %>%
  drop_na(sleepdur_yest, Oura_Sleep_hour)

cor.test(tmp$sleepdur_yest, tmp$Oura_Sleep_hour, method = "spearman")
```

Overall:

```{r fig.height=4, fig.width=12}
# Include all Oura and self-report data points, filter BiAffect data points
# based on train/test label
dat_descr <- dat %>%
  left_join(last_train_days, by = "subject") %>%
  select(subject, date, sleepdur_yest, Oura_Sleep_hour, hour_estimate, 
         label, cleartrialphase, last_train_day) %>%
  pivot_longer(sleepdur_yest:hour_estimate, names_to = "variable") %>%
  filter(
    variable != "hour_estimate" | (label == "test" & date > last_train_day + days(3)),
    !is.na(cleartrialphase)
  )

max_val = round(max(dat_descr$value, na.rm = TRUE))

g_violin <- ggplot(dat_descr, aes(variable, value, fill = variable)) +
  geom_violin() +
  geom_jitter(width = 0.05, height = 0, alpha = 0.1) +
  scale_x_discrete(labels = c("BiAffect", "Oura ring", "Self-report")) +
  scale_y_continuous(minor_breaks = seq(0, max_val)) +
  scale_fill_manual(values = c("lightcoral", "aquamarine3", "skyblue4")) +
  labs(x = NULL, y = "Sleep duration (hour)") +
  # facet_wrap(~ variable, scales = "free_y") +
  theme_minimal() +
  theme(
    text = element_text(size = 16),
    legend.position="none",
    axis.text.x = element_text(size = 16)
  )

g_violin

# ggsave(file.path(man_img_dir, "violin_sleep.pdf"))
# ggsave(file.path(man_img_dir, "gsvd_validation2", "violins3_min30.pdf"))

dat_descr %>%
  drop_na(value) %>%
  group_by(variable) %>%
  summarize(n())
```

```{r}
max_lim <- 19
r_coord = 15

tmp <- dat_descr %>%
  pivot_wider(id_cols = c(subject, date), 
              names_from = variable, values_from = value)

gs <- vector("list", length = 3)

gs[[1]] <- ggplot(tmp, aes(hour_estimate, Oura_Sleep_hour)) +
  labs(x = "BiAffect sleep duration (hour)", 
       y = "Oura sleep duration (hour)") +
  geom_label(aes(x, y), label = "r[s] == 0.40", parse = TRUE, 
             data = data.frame(x = r_coord, y = r_coord)) # Prevent overdrawing

gs[[2]] <- ggplot(tmp, aes(hour_estimate, sleepdur_yest)) +
  labs(x = "BiAffect sleep duration (hour)", 
       y = "Self-reported sleep duration (hour)") +
  geom_label(aes(x, y), label = "r[s] == 0.26", parse = TRUE, 
             data = data.frame(x = r_coord, y = r_coord)) # Prevent overdrawing

gs[[3]] <- ggplot(tmp, aes(sleepdur_yest, Oura_Sleep_hour))+
  labs(x = "Self-reported sleep duration (hour)", 
       y = "Oura sleep duration (hour)") +
  geom_label(aes(x, y), label = "r[s] == 0.74", parse = TRUE, 
             data = data.frame(x = r_coord, y = r_coord)) # Prevent overdrawing
  
gs <- lapply(gs, function(g) {
  g +
    geom_jitter(width = 0.2, height = 0.2, alpha = 0.5) +
    xlim(0, max_lim) +
    ylim(0, max_lim) +
    coord_fixed() +
    theme_minimal() +
    theme(text = element_text(size = 16))
})

g_scatter <- ggarrange(plotlist = gs, nrow = 1)
```

```{r fig.height=8}
ggarrange(g_violin, g_scatter, labels = c("A", "B"), font.label = list(size = 20), nrow = 2)

# ggsave(file.path(man_img_dir, "violin_scatter.pdf"))
```

```{r, fig.height=5, fig.width=10}
tmp <- dat_descr %>%
  filter(variable != "sleepdur_yest")

max_val = round(max(tmp$value, na.rm = TRUE))

g_violin2 <- ggplot(tmp, aes(variable, value, fill = variable)) +
  geom_violin() +
  geom_jitter(width = 0.05, height = 0, alpha = 0.1) +
  scale_x_discrete(labels = c("BiAffect", "Oura ring")) +
  scale_y_continuous(minor_breaks = seq(0, max_val)) +
  scale_fill_manual(values = c("lightcoral", "aquamarine3")) +
  labs(x = NULL, y = "Sleep duration (hour)") +
  # facet_wrap(~ variable, scales = "free_y") +
  theme_minimal() +
  theme(
    text = element_text(size = 32),
    legend.position="none",
    axis.text.x = element_text(size = 32)
  )

g_violin2

# ggsave(file.path(sobp_dir, "violins2.pdf"))
# ggsave(file.path(man_img_dir, "gsvd_validation2", "violins.pdf"))
# ggsave("~/Documents/Presentations/2025-10-06 BYL/violins.png", dpi = 300)
```

```{r fig.height=5, fig.width=11}
ggarrange(g_violin2, gs[[1]], widths = c(2, 2), align = "v", labels = "AUTO")

# ggsave(file.path(man_img_dir, "gsvd_validation2", "violin_scatter.pdf"))
```

```{r fig.height=11, fig.width=7}
ggarrange(
  g_violin2 + theme(text = element_text(size = 20), axis.text.x = element_text(size = 20)), 
  gs[[1]] + theme(text = element_text(size = 20)), 
  ncol = 1, widths = c(2, 2), labels = "AUTO"
)
```


More than enough evidence for random slopes and intercepts:

```{r fig.height=8}
lmList <- nlme::lmList

m_list <- lmList(Oura_Sleep_hour ~ hour_estimate_c | subject, data = dat_test_oura, na.action = na.omit)

plot(intervals(m_list))
```

Ages:

```{r}
tmp <- dat_test_oura %>%
  group_by(subject) %>%
  summarize(Age = Age[1])

hist(tmp$Age, breaks = 30)
```

# Predicting self-report (hour estimate)

This section predicts self-reported sleep duration.

```{r}
cor.test(dat_test_sr$hour_estimate, dat_test_sr$sleepdur_yest, method = "spearman")
```

```{r}
m2_sr <- lmer(sleepdur_yest ~ hour_estimate_c + (hour_estimate_c | subject), 
              dat_test_sr)

summary(m2_sr)
```

To get the p values.

```{r paged.print=FALSE}
m2_nlme <- lme(sleepdur_yest ~ hour_estimate_c, 
               data = dat_test_sr, 
               random = ~ hour_estimate_c | subject)

summary(m2_nlme)
```

```{r}
check_model(m2_sr, sleepdur_yest, hour_estimate_c)
```

## Prediction

```{r fig.height=8}
plot_reg_ev(
  m2_sr,
  dat_test_sr,
  hour_estimate_c,
  hour_estimate,
  sleepdur_yest,
  xlab = "BiAffect-derived sleep duration (hour)",
  ylab = "Self-reported sleep duration (hour)"
)

# ggsave(file.path(man_img_dir, "duration_prediction_sr_var.pdf"))
```

# Predicting time in bed

This section predicts Oura-derived time spent in bed (TIB).

```{r}
cortest <- cor.test(dat_test_oura$hour_estimate, dat_test_oura$TimeInBed_hour,
                    method = "spearman")

cortest
```

```{r}
m2_tib <- lmer(
  TimeInBed_hour ~ hour_estimate_c + (hour_estimate_c | subject),
  dat_test_oura)

summary(m2_tib)
```

```{r paged.print=FALSE}
summary(lme(
  TimeInBed_hour ~ hour_estimate_c,
  random = ~ hour_estimate_c | subject,
  data = dat_test_oura,
  na.action = na.omit
))
```


### Evaluation

```{r}
check_model(m2_tib, TimeInBed_hour, hour_estimate_c)
```

### Prediction

```{r fig.height=8}
plot_reg_ev(
  m2_tib, dat_test_oura, hour_estimate_c, hour_estimate, TimeInBed_hour, 
  xlab = "BiAffect-estimated sleep duration (hour)",
  ylab = "Oura-estimated time in bed (hour)"
)

# ggsave(file.path(man_img_dir, "model_tib.pdf"))
```

# Predicting Oura ring

Here we predict Oura-derived sleep duration using self-report and BiAffect predictors. Further below are sensitivity analyses involving alternative thresholding, filtering, and regularisation methods, as well as some other things.

## Self-report

```{r}
dat_oura_sr <- dat_oura %>%
  mutate(
    sleepdur_yest_c = drop(scale(sleepdur_yest)),
    id = factor(id)
  ) %>%
  drop_na(sleepdur_yest, Oura_Sleep_hour) %>%
  rename(subject = id)

dat_oura_sr
```


```{r}
m2_oura_sr <- lmer(
  Oura_Sleep_hour ~ sleepdur_yest_c + (sleepdur_yest_c | subject),
  dat_oura_sr)

summary(m2_oura_sr)
```

```{r paged.print=FALSE}
summary(
  lme(
    Oura_Sleep_hour ~ sleepdur_yest_c,
    random = ~ sleepdur_yest_c | subject,
    data = dat_oura_sr
  )
)
```

```{r}
rmse_fixed(m2_oura_sr, Oura_Sleep_hour, sleepdur_yest_c)
```

### Evaluation

```{r}
check_model(m2_oura_sr, Oura_Sleep_hour, sleepdur_yest_c)
```

### Prediction

```{r fig.height=8}
plot_reg_ev(
  m2_oura_sr, dat_oura_sr, sleepdur_yest_c, sleepdur_yest, Oura_Sleep_hour, 
  xlab = "Self-reported sleep duration (hour)",
  ylab = "Oura-estimated sleep duration (hour)"
)

# ggsave(file.path(man_img_dir, "model_oura_sr.pdf"))
```



## BiAffect sleep duration

```{r}
cortest <- cor.test(dat_test_oura$hour_estimate, dat_test_oura$Oura_Sleep_hour,
                    method = "spearman")

cortest
```

```{r}
tmp <- dat_test_oura %>%
  mutate(
    sedative_YN = as.integer(sedative_YN),
    stim_YN = as.integer(stim_YN),
    SSRI = as.integer(SSRI),
    SNRI = as.integer(SNRI),
    age_c = drop(age_c)
  )

m2_oura <- lmer(
  # Oura_Sleep_hour ~ hour_estimate_c + age_c + sedative_YN + stim_YN + SSRI + SNRI + (hour_estimate_c | subject),
  Oura_Sleep_hour ~ hour_estimate_c + (hour_estimate_c | subject),
  tmp)

summary(m2_oura)
```

```{r}
tmp <- dat_test_oura %>%
  mutate(hour_estimate_c_unscaled = drop(scale(hour_estimate, scale = FALSE)))

m2_oura_unscaled <- lmer(
  Oura_Sleep_hour ~ hour_estimate_c_unscaled + (hour_estimate_c_unscaled | subject),
  tmp
)

summary(m2_oura_unscaled)
```

```{r}
print(rmse(m2_oura))
print(rmse(m2_oura_unscaled))

print(mae(m2_oura))
print(mae(m2_oura_unscaled))
```


Fixed effects:
                Estimate Std. Error t value
(Intercept)      6.87945    0.13054  52.700
hour_estimate_c  0.72273    0.08586   8.418

Correlation of Fixed Effects:
            (Intr)
hour_stmt_c 0.307 

```{r paged.print=FALSE}
m2_oura_nlme <- lme(Oura_Sleep_hour ~ hour_estimate_c, 
                    data = dat_test_oura, 
                    random = ~ hour_estimate_c | subject)

summary(m2_oura_nlme)
```

```{r}
get_coef_table <- function(m) {
  summ <- summary(m)
  tab <- summ$tTable
  tab <- tab[2:nrow(tab),]
  
  betas <- tab[,1]
  ses <- tab[,2]
  p_vals <- tab[,5]
  
  data.frame(beta = betas, SE = ses, p = p_vals)
}

# get_coef_table(m2_oura_nlme)
```

```{r}
plot(ACF(m2_oura_nlme, resType = "normalized"))
```

```{r paged.print=FALSE}
m2_oura_cor <- lme(Oura_Sleep_hour ~ hour_estimate_c, 
                   data = dat_test_oura, 
                   random = ~ hour_estimate_c | subject,
                   correlation = corCAR1(form = ~ date | subject))

summary(m2_oura_cor)
```


### With covariates

```{r paged.print=FALSE}
tmp <- dat_test_oura %>%
  mutate(
    sedative_YN = as.integer(sedative_YN),
    stim_YN = as.integer(stim_YN),
    SSRI = as.integer(SSRI),
    SNRI = as.integer(SNRI),
    age_c = drop(age_c)
  )

m2_oura_cov <- lme(
  Oura_Sleep_hour ~ hour_estimate_c + age_c + sedative_YN + stim_YN + SSRI,
  random = ~ hour_estimate_c | subject,
  data = tmp,
  na.action = na.omit)

summary(m2_oura_cov)
```

```{r}
r2mlm(m2_oura_cov)
```

### Cluster-centered

Cluster-centred predictors.

```{r}
cluster_lms <- nlme::lmList(Oura_Sleep_hour ~ hour_estimate_cluster_centered + hour_estimate_cluster_mean | subject,
                            data = dat_test_oura, na.action = na.omit)

plot(intervals(cluster_lms))
```

lmer complains here, but the lme4 troubleshooting page suggests "refitting the parameters starting at the current estimates: getting consistent results (with no warning) suggests a false positive". Doing such a refit eliminates the warning and returns almost identical results, so I think these results are fine.

I could alternatively scale hour_estimate_cluster_centered within participants. That fixes the warning, but reduces the explained variance. I suppose scalin within participants throws away information? Not sure what the best thing to do would be.

Scaling hour_estimate_cluster_centered across all participants (so first center within participants, then center again AND scale to unit variance across the entire sample) also fixes the lmer warning. I think this might be the best option.

```{r}
m2_oura_cluster <- lmer(
  Oura_Sleep_hour ~ hour_estimate_cluster_centered + hour_estimate_cluster_mean + (hour_estimate_cluster_centered | subject),
  dat_test_oura)

summary(m2_oura_cluster)
```

```{r}
m_tmp <- lmer(
  Oura_Sleep_hour ~ hour_estimate_cluster_centered + hour_estimate_cluster_mean + (hour_estimate_cluster_centered | subject),
  dat_test_oura, start = m2_oura_cluster@theta)

summary(m_tmp)
```

```{r}
r2mlm(m2_oura_cluster)
```

```{r fig.height=8}
plot(predict_response(m2_oura_cluster, terms = "hour_estimate_cluster_mean"), show_data = TRUE, jitter = 0.01) +
  labs(title = "Predicted values of sleep duration", 
       x = "Hour estimate (between)", y = "Sleep duration") +
  theme(text = element_text(size = 16))

plot(predict_response(m2_oura_cluster, terms = "hour_estimate_cluster_centered"), show_data = TRUE, jitter = 0.01) +
  labs(title = "Predicted values of sleep duration", 
       x = "Hour estimate (within)", y = "Sleep duration") +
  theme(text = element_text(size = 16))
```

### Evaluation

```{r}
resp <- m2_oura@frame$Oura_Sleep_hour
sqrt(mean((resp - mean(resp))^2))
mean(abs(resp - mean(resp)))
```


```{r}
check_model(m2_oura, Oura_Sleep_hour, hour_estimate_c)
```

```{r}
hist(get_raw_resids(m2_oura, Oura_Sleep_hour, hour_estimate_c), breaks = 30)
```


Random intercepts look okay, random slopes show some structure.

```{r}
ls <- plot_ev(m2_oura, bargraph = TRUE)

g_var <- ls$g_var
g_var_hor <- ls$g_var_hor

ls
```


```{r}
r2mlm(m2_oura_cluster)
```

```{r}
anova(m2_oura_nlme)
```

F-test, fixed effects

```{r}
anova(m2_oura_nlme)
```

### Baseline comparison

Comparison to baseline model, fixed effects

```{r paged.print=FALSE}
m2_oura_baseline <- update(
  m2_oura_nlme, 
  fixed = Oura_Sleep_hour ~ 1, 
  random = ~ 1 | subject, 
  method = "ML"
)

m2_oura_fixed <- update(
  m2_oura_nlme,
  random = ~ 1 | subject,
  method = "ML"
)

anova(m2_oura_baseline, m2_oura_fixed)
```

```{r}
print(rmse(m2_oura_baseline))
print(rmse(m2_oura_fixed))

cat("\n")
print(mae(m2_oura_baseline))
print(mae(m2_oura_fixed))
```

For df=1, the nominal and empirical p-values match up, which is good (this is the df we use for anova). For higher dfs, the nominal p-values tend to be higher than the empirical values, which would indicate a conservative test.

```{r}
lme_sim <- simulate.lme(
  list(
    fixed = Oura_Sleep_hour ~ 1, 
    random = ~ 1 | subject, 
    data = groupedData(Oura_Sleep_hour ~ hour_estimate_c | subject, dat_test_oura)
  ),
  m2 = list(fixed = Oura_Sleep_hour ~ hour_estimate_c),
  method = "ML", nsim = 1000
)

plot(lme_sim, df = c(1, 2))
```

Comparison to baseline model, random effects

```{r paged.print=FALSE}
m2_oura_baseline2 <- update(
  m2_oura_nlme, 
  random = ~ 1 | subject
)

anova(m2_oura_baseline2, m2_oura_nlme)
```

For df=2, the test would be a bit conservative.

```{r}
lme_sim <- simulate.lme(
  list(
    fixed = Oura_Sleep_hour ~ hour_estimate_c, 
    random = ~ 1 | subject, 
    data = groupedData(Oura_Sleep_hour ~ hour_estimate_c | subject, dat_test_oura)
  ),
  m2 = list(random = ~ hour_estimate_c | subject),
  method = "REML", nsim = 1000
)

plot(lme_sim, df = c(1, 2, 3))
```

```{r}
print(rmse(m2_oura_baseline2))
print(rmse(m2_oura_nlme))

cat("\n")
print(mae(m2_oura_baseline2))
print(mae(m2_oura_nlme))
```

### Prediction

```{r fig.height=8}
plot(predict_response(m2_oura, terms = "hour_estimate_c"), show_data = TRUE, jitter = 0.1) +
  labs(title = "Predicted values of sleep duration", 
       x = "Hour estimate (centred)", y = "Sleep duration") +
  theme(text = element_text(size = 16))
```

```{r fig.height=8}
g_pred <- plot_regression(
  m2_oura, dat_test_oura, hour_estimate_c, hour_estimate, Oura_Sleep_hour,
  xlab = "BiAffect-estimated sleep duration (hour)",
  ylab = "Oura-estimated sleep duration (hour)"
)

g_pred

# ggsave(file.path(man_img_dir, "duration_prediction_oura.pdf"))
```

```{r fig.height=8, fig.width=12}
plot_reg_ev(
  m2_oura, dat_test_oura, hour_estimate_c, hour_estimate, Oura_Sleep_hour,
  xlab = "BiAffect-estimated sleep duration (hour)",
  ylab = "Oura-estimated sleep duration (hour)"
)

# ggsave(file.path(man_img_dir, "gsvd_validation2", "duration_prediction_oura_var_min30.pdf"))
```

```{r fig.height=8, fig.width=12}
ggarrange(g_pred, g_var, widths = c(6, 1), legend.grob = get_legend(g_var_hor))
# 
# ggsave(file.path(sobp_dir, "duration_prediction_oura_var.pdf"))
```


## Threshold sensitivity analysis

### Otsu's threshold

```{r}
m2_otsu <- lmer(
  Oura_Sleep_hour ~ hour_estimate_otsu_c + (hour_estimate_otsu_c | subject),
  dat_test_oura)

summary(m2_otsu)
```

```{r paged.print=FALSE}
summary(lme(
  Oura_Sleep_hour ~ hour_estimate_otsu_c,
  random = ~ hour_estimate_otsu_c | subject,
  data = dat_test_oura,
  na.action = na.omit
))
```


```{r}
check_model(m2_otsu, Oura_Sleep_hour, hour_estimate_otsu_c)
```

```{r fig.height=8}
plot_reg_ev(m2_otsu, dat_test_oura, hour_estimate_otsu_c, hour_estimate_otsu, Oura_Sleep_hour, 
            xlab = "BiAffect-derived sleep duration", ylab = "Ring-derived sleep duration")

# ggsave(file.path(man_img_dir, "model_otsu.pdf"))
```

### Gaussian mixture

```{r}
m2_gm <- lmer(
  Oura_Sleep_hour ~ hour_estimate_gm_c + (hour_estimate_gm_c | subject),
  dat_test_oura)

summary(m2_gm)
```

```{r paged.print=FALSE}
summary(lme(
  Oura_Sleep_hour ~ hour_estimate_gm_c,
  random = ~ hour_estimate_gm_c | subject,
  data = dat_test_oura,
  na.action = na.omit
))
```

```{r}
check_model(m2_gm, Oura_Sleep_hour, hour_estimate_gm_c)
```

```{r fig.height=8}
plot_reg_ev(m2_gm, dat_test_oura, hour_estimate_gm_c, hour_estimate_gm, Oura_Sleep_hour, 
            xlab = "BiAffect-derived sleep duration", ylab = "Ring-derived sleep duration")
4
# ggsave(file.path(man_img_dir, "model_gm.pdf"))
```

## Data exclusion sensitivity analysis

### 16 percent (4 hours)

```{r echo=TRUE, results='hide'}
dat_test_16 <- sleep_preds_svd(
  file.path(
    dat_dir, 
    "svd_mats_long_alpha1_2025-08-19_utc-reference_min30_16percent.parquet"
  ),
  dat_dir,
  dat_oura
) %>%
  left_join(dem, by = c("subject" = "id")) %>%
  drop_na(c(hour_estimate, Oura_Sleep_hour)) %>%
  filter_mutate()

dat_test_16
```

```{r}
hist(dat_test_16$hour_estimate, breaks = 24)
```

```{r}
m2_16 <- lmer(
  Oura_Sleep_hour ~ hour_estimate_c + (hour_estimate_c | subject),
  dat_test_16)

summary(m2_16)
```

```{r paged.print=FALSE}
summary(lme(
  Oura_Sleep_hour ~ hour_estimate_c,
  random = ~ hour_estimate_c | subject,
  data = dat_test_16,
  na.action = na.omit
))
```

#### Evaluation

```{r}
check_model(m2_16, Oura_Sleep_hour, hour_estimate_c)
```

#### Prediction

```{r fig.height=8}
plot_reg_ev(
  m2_16, dat_test_16, hour_estimate_c, hour_estimate, Oura_Sleep_hour, 
  xlab = "BiAffect-estimated sleep duration (hour)",
  ylab = "Oura-estimated sleep duration (hour)"
)

# ggsave(file.path(man_img_dir, "model_16percent.pdf"))
```

### 50 percent

```{r echo=TRUE, results='hide'}
dat_test_50 <- sleep_preds_svd(
  file.path(
    dat_dir, 
    "svd_mats_long_alpha1_2025-08-19_utc-reference_min30_50percent.parquet"
  ),
  dat_dir,
  dat_oura
) %>%
  left_join(dem, by = c("subject" = "id")) %>%
  drop_na(c(hour_estimate, Oura_Sleep_hour)) %>%
  filter_mutate()

dat_test_50
```

```{r}
m2_50 <- lmer(
  Oura_Sleep_hour ~ hour_estimate_c + (hour_estimate_c | subject),
  dat_test_50)

summary(m2_50)
```

```{r paged.print=FALSE}
summary(lme(
  Oura_Sleep_hour ~ hour_estimate_c,
  random = ~ hour_estimate_c | subject,
  data = dat_test_50,
  na.action = na.omit
))
```

#### Evaluation

```{r}
check_model(m2_50, Oura_Sleep_hour, hour_estimate_c)
```


```{r}
plot(m2_50)
```

```{r}
qqnorm(resid(m2_50))
qqline(resid(m2_50))

hist(resid(m2_50), breaks = 30)
```

```{r}
re <- ranef(m2_50)$subject %>%
  rownames_to_column("subject") %>%
  pivot_longer(!subject)

ggplot(re, aes(sample = value)) +
  geom_qq() +
  geom_qq_line() +
  facet_wrap(~ name, scales = "free")
```

```{r}
r2mlm(m2_50)
rmse(m2_50)
mae(m2_50)
```

#### Prediction

```{r fig.height=8}
plot_reg_ev(
  m2_50, dat_test_50, hour_estimate_c, hour_estimate, Oura_Sleep_hour, 
  xlab = "BiAffect-estimated sleep duration (hour)",
  ylab = "Oura-estimated sleep duration (hour)"
)

# ggsave(file.path(man_img_dir, "model_50percent.pdf"))
```

### 66 percent

```{r echo=TRUE, results='hide'}
dat_test_66 <- sleep_preds_svd(
  file.path(
    dat_dir, 
    "svd_mats_long_alpha1_2025-08-19_utc-reference_min30_66percent.parquet"
  ),
  dat_dir,
  dat_oura
) %>%
  left_join(dem, by = c("subject" = "id")) %>%
  drop_na(c(hour_estimate, Oura_Sleep_hour)) %>%
  filter_mutate()

dat_test_66
```

```{r}
m2_66 <- lmer(
  Oura_Sleep_hour ~ hour_estimate_c + (hour_estimate_c | subject),
  dat_test_66)

summary(m2_66)
```

#### Evaluation

```{r}
check_model(m2_66, Oura_Sleep_hour, hour_estimate_c)
```

#### Prediction

```{r fig.height=8}
plot_reg_ev(
  m2_66, dat_test_66, hour_estimate_c, hour_estimate, Oura_Sleep_hour, 
  xlab = "BiAffect-estimated sleep duration (hour)",
  ylab = "Oura-estimated sleep duration (hour)"
)

# ggsave(file.path(man_img_dir, "model_50percent.pdf"))
```

## Regularisation sensitivity analysis

### λ = 0.1

```{r echo=TRUE, results='hide'}
dat_test_alpha01 <- sleep_preds_svd(
  file.path(
    dat_dir, 
    "svd_mats_long_alpha0.1_2025-08-19_utc-reference_min30.parquet"
  ),
  dat_dir,
  dat_oura
) %>%
  left_join(dem, by = c("subject" = "id")) %>%
  drop_na(c(hour_estimate, Oura_Sleep_hour)) %>%
  filter_mutate()

dat_test_alpha01
```

```{r}
m2_alpha01 <- lmer(
  Oura_Sleep_hour ~ hour_estimate_c + (hour_estimate_c | subject),
  dat_test_alpha01)

summary(m2_alpha01)
```

```{r paged.print=FALSE}
summary(
  lme(
    Oura_Sleep_hour ~ hour_estimate_c,
    random = ~ hour_estimate_c | subject,
    data = dat_test_alpha01
  )
)
```


#### Evaluation

```{r}
check_model(m2_alpha01, Oura_Sleep_hour, hour_estimate_c)
```

#### Prediction

```{r fig.height=8}
plot_reg_ev(
  m2_alpha01, dat_test_alpha01, hour_estimate_c, hour_estimate, Oura_Sleep_hour, 
  xlab = "BiAffect-estimated sleep duration (hour)",
  ylab = "Oura-estimated sleep duration (hour)"
)

# ggsave(file.path(man_img_dir, "model_alpha01.pdf"))
```


### λ = 10

```{r echo=TRUE, results='hide'}
dat_test_alpha10 <- sleep_preds_svd(
  file.path(
    dat_dir, 
    "svd_mats_long_alpha10_2025-08-19_utc-reference_min30.parquet"
  ),
  dat_dir,
  dat_oura,
  sleep_threshold = 1.25
) %>%
  left_join(dem, by = c("subject" = "id")) %>%
  drop_na(c(hour_estimate, Oura_Sleep_hour)) %>%
  filter_mutate()

dat_test_alpha10
```

```{r}
m2_alpha10 <- lmer(
  Oura_Sleep_hour ~ hour_estimate_c + (hour_estimate_c | subject),
  dat_test_alpha10)

summary(m2_alpha10)
```

```{r paged.print=FALSE}
summary(
  lme(
    Oura_Sleep_hour ~ hour_estimate_c,
    random = ~ hour_estimate_c | subject,
    data = dat_test_alpha10
  )
)
```

#### Evaluation

```{r}
check_model(m2_alpha10, Oura_Sleep_hour, hour_estimate_c)
```

#### Prediction

```{r fig.height=8}
plot_reg_ev(
  m2_alpha10, dat_test_alpha10, hour_estimate_c, hour_estimate, Oura_Sleep_hour, 
  xlab = "BiAffect-estimated sleep duration (hour)",
  ylab = "Oura-estimated sleep duration (hour)"
)

# ggsave(file.path(man_img_dir, "model_alpha10.pdf"))
```

## Multimodality sensitivity analysis

```{r echo=TRUE, results='hide'}
dat_test_mm <- sleep_preds_svd(
  file.path(
    dat_dir, 
    "svd_mats_long_alpha1_2025-08-22_extra_modalities.parquet"
  ),
  dat_dir,
  dat_oura,
  sleep_threshold = 1.5
) %>%
  left_join(dem, by = c("subject" = "id")) %>%
  drop_na(c(hour_estimate, Oura_Sleep_hour)) %>%
  filter_mutate()

dat_test_mm
```

```{r}
hist(dat_test_mm$hour_estimate, breaks = 24)
```


```{r}
m2_mm <- lmer(
  Oura_Sleep_hour ~ hour_estimate_c + (hour_estimate_c | subject),
  dat_test_mm)

summary(m2_mm)
```

```{r paged.print=FALSE}
summary(
  lme(
    Oura_Sleep_hour ~ hour_estimate_c,
    random = ~ hour_estimate_c | subject,
    data = dat_test_mm
  )
)
```

### Evaluation

```{r}
check_model(m2_mm, Oura_Sleep_hour, hour_estimate_c)
```

### Prediction

```{r fig.height=8}
plot_reg_ev(
  m2_mm, dat_test_mm, hour_estimate_c, hour_estimate, Oura_Sleep_hour, 
  xlab = "BiAffect-estimated sleep duration (hour)",
  ylab = "Oura-estimated sleep duration (hour)"
)

# ggsave(file.path(man_img_dir, "model_mm.pdf"))
```

## Clipping sensitivity analysis

Only a single data point is being clipped.

```{r}
clip_val <- function(val, mini, maxi) pmin(maxi, pmax(mini, val))

dat_test_clip <- dat_test_oura %>%
  mutate(
    hour_estimate_clipped = clip_val(hour_estimate, 1, 13),
    hour_estimate_clipped_c = drop(scale(hour_estimate_clipped))
  )
```

```{r}
m2_clip <- lmer(
  Oura_Sleep_hour ~ hour_estimate_clipped_c + (hour_estimate_clipped_c | subject),
  dat_test_clip)

summary(m2_clip)
```

```{r paged.print=FALSE}
summary(
  lme(
    Oura_Sleep_hour ~ hour_estimate_clipped_c,
    random = ~ hour_estimate_clipped_c | subject,
    data = dat_test_clip
  )
)
```

#### Evaluation

```{r}
check_model(m2_clip, Oura_Sleep_hour, hour_estimate_clipped_c)
```

#### Prediction

```{r fig.height=8}
plot_reg_ev(
  m2_clip, dat_test_clip, hour_estimate_clipped_c, hour_estimate_clipped, Oura_Sleep_hour, 
  xlab = "Clipped BiAffect-estimated sleep duration (hour)",
  ylab = "Oura-estimated sleep duration (hour)"
)

# ggsave(file.path(man_img_dir, "model_clipped.pdf"))
```

# Demographics printing

This formats the demographics for the manuscript.

```{r}
ids <- as.integer(unique(dat_test_sr$subject))

dem_test <- dem %>%
  filter(id %in% ids) %>%
  select(!c(id, race_clear123, ethnicity, bmi, filing_status, sex_orient)) %>%
  select(where(~ !is.numeric(.x) || sum(.x, na.rm = TRUE) > 0)) %>%
  mutate(across(starts_with("currentdx"), as.logical))

tab_txt <- capture.output(CreateTableOne(data = dem_test)) %>%
  str_replace("^\\s\\s", "") %>%
  str_replace("(\\w|\\))\\s+(\\d.*)", "\\1;\\2") %>%
  str_replace_all("currentdx_(.+) = TRUE \\(%\\);(.*)", "   \\1;\\2")

cat(tab_txt, sep = "\n")
```

```{r}
ids <- as.integer(unique(dat_test_oura$subject))

dem_test <- dem %>%
  filter(id %in% ids) %>%
  select(!c(id, race_clear123, ethnicity, bmi, filing_status, sex_orient)) %>%
  select(where(~ !is.numeric(.x) || sum(.x, na.rm = TRUE) > 0)) %>%
  mutate(across(starts_with("currentdx"), as.logical))

tab_txt <- capture.output(CreateTableOne(data = dem_test)) %>%
  str_replace("^\\s\\s", "") %>%
  str_replace("(\\w|\\))\\s+(\\d.*)", "\\1;\\2") %>%
  str_replace_all("currentdx_(.+) = TRUE \\(%\\);(.*)", "   \\1;\\2")

cat(tab_txt, sep = "\n")
```

```{r}
ids <- as.integer(unique(dat_test_sr$subject))

t1_sr <- create_tab1(dem, ids)

ids <- as.integer(unique(dat_test_oura$subject))

t1_oura <- create_tab1(dem, ids)

ids <- readRDS(file.path(dat_dir, "ids_phase_analysis.rds"))

t1_phase <- create_tab1(dem, ids)

dem_tab <- t1_sr %>%
  full_join(t1_oura, by = c("variable", "__var", "type", "order")) %>%
  rename(`Self-report` = text.x, `Oura ring` = text.y) %>%
  full_join(t1_phase, by = c("variable", "__var", "type", "order")) %>%
  rename(`Phase analysis` = text) %>%
  arrange(order, `__var`) %>%
  select(!c(type, `__var`, order)) %>%
  mutate(across(everything(), ~ replace_na(.x, "")))

write_csv(dem_tab, "~/Documents/Writing/Papers/Paper 2 GSVD/table1v2.csv")

dem_tab
```

# Druijff-Van de Woestijne

This model is based on an alternative method to predict sleep from typing data.

```{r}
dat_dvdw <- read_parquet(file.path(dat_dir, "dat_dvdw_2025-09-01.parquet"))

colnames(dat_dvdw)
```

```{r}
dat_dvdw_oura <- dat_dvdw %>%
  left_join(dat_oura %>% mutate(id = as.character(id)), 
            by = c("subject" = "id", "date" = "daterated")) %>%
  drop_na(Oura_Sleep_TST) %>%
  mutate(
    kap_c = drop(scale(kap)),
    n_active_hours_c = drop(scale(n_active_hours)),
    prev_n_active_hours_c = drop(scale(prev_n_active_hours))
  )

colnames(dat_dvdw_oura)
```

```{r}
m_dvdw <- lmer(Oura_Sleep_hour ~ kap_c + n_active_hours_c + prev_n_active_hours_c + (1 | subject),
               dat_dvdw_oura)

summary(m_dvdw)
```

```{r paged.print=FALSE}
m_dvdw_nlme <- lme(Oura_Sleep_hour ~ kap_c + n_active_hours_c + prev_n_active_hours_c, 
                   data = dat_dvdw_oura,
                   random = ~ 1 | subject,
                   na.action = na.omit)

summary(m_dvdw_nlme)
```


```{r}
plot(m_dvdw)

qqnorm(resid(m_dvdw))
qqline(resid(m_dvdw))

hist(resid(m_dvdw))

re <- ranef(m_dvdw)$subject %>%
  rownames_to_column("subject") %>%
  pivot_longer(!subject)

ggplot(re, aes(sample = value)) +
  geom_qq() +
  geom_qq_line() +
  facet_wrap(~ name, scales = "free")
```

```{r}
plot(predict_response(m_dvdw, terms = "kap_c"), show_data = TRUE) +
  xlim(-4, 4)
```


```{r fig.height=8}
kap_mean <- mean(dat_dvdw_oura$kap)
kap_sd <- sd(dat_dvdw_oura$kap)

pr <- predict_response(m_dvdw, terms = "kap_c") %>%
  as.data.frame() %>%
  mutate(x = kap_sd * x + kap_mean)

n_presses_breaks = c(400, 1100, 3000, 8100, 22000)

g_dvdw <- ggplot(pr, aes(x, predicted)) +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "grey") +
  geom_line(color = "lightcoral", lwd = 1.2) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), fill="lightcoral", alpha = 0.2) +
  geom_point(
    aes(kap, Oura_Sleep_hour),
    data = dat_dvdw_oura) +
  # scale_color_gradient(name = "Number of key presses", transform = "log",
  #                      breaks = n_presses_breaks, labels = n_presses_breaks, high = "aquamarine3") +
  scale_x_continuous(minor_breaks = ~ seq(0, 15)) +
  labs(x = "Keystroke-absence period (hour)", y = "Oura-estimated sleep duration (hour)") +
  theme_minimal() +
  theme(
    text = element_text(size = 20)
  )

# ggsave(file.path(man_img_dir, "duration_prediction_dvdw.pdf"))

g_dvdw
```

```{r}
r_squared <- r2mlm(m_dvdw)

# noquote class
vars <- r_squared$Decompositions[,1]
vars_df <- data.frame(variance = unclass(vars)) %>%
  rownames_to_column("component") %>%
  filter(component != "fixed, between", component != "slope variation") %>%
  mutate(
    component = case_when(
      component == "fixed" ~ "Fixed effects",
      component == "mean variation" ~ "Random intercept",
      component == "sigma2" ~ "Residuals"
    )
  )

vars_df

cols <- rev(c("aquamarine2", "aquamarine3", "grey20"))

g_var_dvdw <- ggplot(vars_df, aes(1, variance)) +
  geom_col(aes(fill = fct_rev(component))) +
  scale_fill_manual(values = cols, name = "Variance component") +
  labs(x = NULL, y = "Proportion of variance") +
  scale_x_continuous(breaks = NULL) +
  scale_y_continuous(position = "right") +
  theme_minimal() +
  theme(
    text = element_text(size = 20),
    legend.title = element_text(size = 16),
    legend.text = element_text(size = 14)
  )

g_var_dvdw
```

```{r fig.height=8}
ggarrange(g_dvdw, g_var_dvdw, widths = c(5, 2))

# ggsave(file.path(man_img_dir, "duration_prediction_oura_var_dvdw_2025-09-01.pdf"))
```

# Fluctuations

This is some code to check visually if there might be a relationship between typing activity rhythm disturbance and a poor night's sleep.

```{r echo=TRUE, results='hide'}
svd_long <- read_parquet(file.path(dat_dir, "svd_mats_long_alpha1_2024-11-27.parquet")) %>%
  mutate(
    timestamp = with_tz(timestamp, tzone = "UTC"),
    timestamp_warped = with_tz(timestamp_warped, tzone = "UTC")
  )

svd_long
```

```{r fig.height=8}
point_date <- ymd("2022-08-14") # c(ymd("2022-08-18"), ymd("2022-09-29"))

participant <- "3094" # "3086"

ggplot(svd_long %>% filter(subject == participant), 
       aes(hour(timestamp), as_date(timestamp))) +
  geom_raster(aes(fill = svd_value)) +
  geom_point(aes(x, y), data = data.frame(x = -1, y = point_date), 
             color = "red", size = 4) +
  scale_fill_continuous(type = "viridis") +
  scale_y_continuous(transform = c("date", "reverse")) +
  labs(x = "Hour of the day", y = "Date", fill = "Typing activity") +
  theme(text = element_text(size = 18))
```

```{r fig.height=8}
ggplot(svd_long %>% filter(subject == participant), aes(timestamp, svd_value)) +
  geom_tile(aes(y = 0, fill = label, height = 0.05), alpha = 0.5) +
  geom_point(aes(color = svd_value)) +
  geom_line() +
  geom_point(aes(x, y), data = data.frame(x = as_datetime(point_date), y = 0), 
             color = "red", size = 4) +
  scale_color_continuous(type = "viridis") +
  labs(x = "Time", y = "Typing activity", color = "Typing activity", fill = "Train/test label") +
  theme(text = element_text(size = 18))
```

# SVD matrix poster SOBP

This prints some graphics for a poster Loran presented at SOBP 2025.

```{r results='hide'}
svd_sobp <- read_parquet(file.path(dat_dir, "sobp_df_3009_alpha1.parquet")) %>%
  mutate(
    timestamp = with_tz(timestamp, tzone = "UTC")
  )

svd_sobp
```

```{r fig.width=8, fig.height=12}
sobp_dir <- "~/Documents/Presentations/Posters/2025-04-26 SOBP/"

ggplot(svd_sobp, 
       aes(hour(timestamp), as_date(timestamp))) +
  geom_tile(aes(fill = svd_value)) +
  scale_fill_continuous(type = "viridis") +
  scale_x_continuous(expand=c(0.01,0)) +
  scale_y_continuous(transform = c("date", "reverse"), expand=c(0.01,0)) +
  labs(x = "Hour of the day", y = "Date", fill = "Typing activity: ") +
  theme_minimal() +
  theme(
    text = element_text(size = 40),
    legend.position = "top",
    legend.key.width = unit(0.5, "inches")
  )

# ggsave(file.path(sobp_dir, "svd_mat.pdf"))
```

```{r fig.height=11, fig.width=6}
ggplot(svd_sobp, 
       aes(hour(timestamp), as_date(timestamp))) +
  geom_tile(aes(fill = svd_value > 1)) +
  scale_fill_discrete(type = c("#4c72b0", "#dd8452")) +
  scale_x_continuous(expand = expansion()) +
  scale_y_continuous(transform = c("date", "reverse"), expand = expansion()) +
  labs(x = "Hour of the day", y = NULL) +
  theme_minimal() +
  theme(
    text = element_text(size = 40),
    legend.position = "none",
    axis.text.y = element_blank()
  )

# ggsave(file.path(sobp_dir, "binarised.pdf"))
```

```{r fig.height=11, fig.width=4}
tmp <- svd_sobp %>%
  mutate(
    sleep = svd_value < 1,
    date = as_date(date)
  ) %>%
  group_by(date) %>%
  summarize(n_sleep = sum(sleep))

ggplot(tmp, aes(n_sleep, date)) +
  geom_line(orientation = "y") +
  geom_point() +
  scale_x_continuous(expand = expansion()) +
  scale_y_continuous(transform = c("date", "reverse"), expand = expansion()) +
  labs(x = "Sleep count", y = NULL) +
  coord_cartesian(clip = "off") +
  theme_minimal() +
  theme(
    text = element_text(size = 40),
    axis.text.y = element_blank()
  )

# ggsave(file.path(sobp_dir, "sleep_count.pdf"))
```



