---
title: "Sleep duration regression CLEAR-3"
output:
  html_document:
    toc: yes
    df_print: paged
  pdf_document:
    toc: yes
  html_notebook:
    toc: yes
    df_print: paged
    toc_float: yes
    theme: spacelab
---

```{r setup}
library(haven)
library(MuMIn)
library(arrow)
library(tidyverse)
library(lme4)
library(nlme)
library(ggeffects)
library(rlang)
library(tableone)
library(r2mlm)
library(ggpubr)

# Read environment variables from a .env file
readRenviron(".env")

dat_dir <- Sys.getenv("DAT_DIR") # Data directory
man_img_dir <- Sys.getenv("MAN_IMG_DIR") # Manuscript image directory
```

# Demographics

```{r}
dem <- readRDS(file.path(dat_dir, "CLEAR3_Biaffect_Leow", "clear3trait.rds")) %>%
  as_factor() %>% # Convert haven_labelled variables to factors
  mutate(
    id = as.character(id),
    race_clear123 = as_factor(race_clear123),
    ethnicity = as_factor(ethnicity)
  ) %>%
  arrange(id) %>%
  mutate(across(where(is.character), ~ if_else(.x == "", "Unknown", .x))) %>%
  rename(Age = age, Gender = gender, `House income` = houseincomecat)

# Given to me by Anisha
dem[dem$id == "3085",]$Age <- 24
dem[dem$id == "3100",]$Age <- 23

colnames(dem)
```

# Oura ring

```{r}
dat_oura <- readRDS(
  file.path(dat_dir, "CLEAR3_Biaffect_Leow", "clear3baseline_sleep.rds")) %>%
  mutate(
    Oura_Sleep_hour = Oura_Sleep_TST / 3600, # Second to hour
    sleepdur_yest = sleepdur_yest + 1 # Likert to hours
  ) %>%
  filter(id != 3086 | daterated != ymd("2022-07-31")) # Duplicate entries

colnames(dat_oura)
```

Convert to parquet to make accessible to Python.

```{r}
dat_oura_raw <- readRDS(
  file.path(dat_dir, "CLEAR3_Biaffect_Leow", "clear3baseline_sleep.rds"))

write_parquet(dat_oura_raw, file.path(dat_dir, 
                                      "CLEAR3_Biaffect_Leow", 
                                      "clear3baseline_sleep.parquet"))
```


```{r fig.height=10}
ggplot(dat_oura, aes(sleepdur_yest, Oura_Sleep_hour)) +
  geom_jitter() +
  geom_abline(intercept = 0, slope = 1) +
  labs(title = "Recorded vs reported sleep time across participants (jittered)") +
  xlab("Self-reported sleep duration (hour)") +
  ylab("Oura-recorded total sleep time (hour)") +
  theme(text = element_text(size = 18))

ggplot(dat_oura, aes(SleepLNQuality, Oura_Sleep_hour)) +
  geom_point() +
  labs(title = "Recorded sleep time vs reported sleep quality across participants") +
  xlab("Self-reported sleep quality (higher is better)") +
  ylab("Oura-recorded total sleep time (hour)") +
  theme(text = element_text(size = 18))

ggplot(dat_oura, aes(sleepdur_yest, SleepLNQuality)) +
  geom_jitter() +
  labs(title = "Reported sleep time vs quality across participants") +
  xlab("Self-reported sleep duration (hour)") +
  ylab("Self-reported sleep quality (higher is better)") +
  theme(text = element_text(size = 18))
```

Participants with self-report data:

```{r}
tmp <- dat_oura %>%
  drop_na(sleepdur_yest)

n_distinct(tmp$id)
```

Participants with Oura data:

```{r}
tmp <- dat_oura %>%
  drop_na(Oura_Sleep_TST)

n_distinct(tmp$id)
```

```{r}
cor.test(dat_oura$sleepdur_yest, dat_oura$Oura_Sleep_hour, method = "spearman")
```

# Reading data

```{r}
# The values in sleepdur_yest are not the raw number of hours slept, but
# rather an index to the response options:
# 1 - 2 or fewer hours
# 2 - 3 hours
# 3 - 4 hours
# ...
# 9 - 10 or more hours
dat <- read_parquet(file.path(dat_dir, "spleep_pred_10_90_alpha1_2024-11-14.parquet")) %>%
  mutate(
    Oura_Sleep_hour = Oura_Sleep_TST / 3600, # Second to hour
    sleepdur_yest = sleepdur_yest + 1
  ) %>%
  filter(subject != "3086" | date != ymd("2022-07-31")) # Duplicate entries

dat_dem <- dat %>%
  left_join(dem, by = c("subject" = "id"))

dat_cmplt_sr <- dat_dem %>%
  drop_na(hour_estimate, sleepdur_yest)

dat_cmplt_oura <- dat_dem %>%
  drop_na(c(hour_estimate, Oura_Sleep_hour))

colnames(dat)
```

Exclude the first few days (amounts to a total of 115) to prevent leakage of information from the test set via temporal correlation.
44 participants.

```{r}
# We can only reliably find the last training days in the full data set
last_train_days <- dat %>%
  drop_na(dayNumber) %>%
  filter(label == "train") %>%
  group_by(subject) %>%
  summarize(last_train_day = tail(dayNumber, n = 1))

# Helper function
filter_mutate <- function(.x) {
  .x %>%
    left_join(last_train_days, by = "subject") %>%
    filter(label == "test", dayNumber > last_train_day + 3) %>% # Don't include first three days in test region
    mutate(
      activity_c = scale(activity),
      hour_estimate_c = drop(scale(hour_estimate)),
      age_c = scale(Age)
    ) %>%
    group_by(subject) %>%
    mutate(
      hour_estimate_cluster_centered = drop(scale(hour_estimate, scale = FALSE)),
      hour_estimate_cluster_mean = mean(hour_estimate, na.rm = TRUE)
    ) %>%
    ungroup()
}

dat_test_sr <- dat_cmplt_sr %>%
  filter_mutate()

dat_test_oura <- dat_cmplt_oura %>%
  filter_mutate()

colnames(dat_test_oura)
```

# Descriptive statistics

```{r fig.height=8, fig.width=8}
ggplot(dat_oura %>% drop_na(Oura_Sleep_TST), aes(id)) +
  geom_bar() +
  labs(x = "Participant ID", y = "Count") +
  theme(
    text = element_text(size = 18),
    axis.text.x = element_text(angle = 90, vjust = 0.5)
  )

ggplot(dat_test_oura, aes(subject)) +
  geom_bar() +
  labs(x = "Participant ID", y = "Count") +
  theme(
    text = element_text(size = 18),
    axis.text.x = element_text(angle = 90, vjust = 0.5)
  )
```

Self-report.

```{r fig.height=8}
ggplot(dat_test_sr, aes(sleepdur_yest)) +
  geom_histogram() +
  labs(x = "Self-reported sleep duration")

ggplot(dat_test_sr, aes(SleepLNQuality)) +
  geom_histogram() +
  labs(x = "Self-reported sleep quality")

ggplot(dat_test_sr, aes(activity)) +
  geom_histogram() +
  labs(x = "Activity score")

ggplot(dat_test_sr, aes(hour_estimate)) +
  geom_histogram()

ggplot(dat_test_sr, aes(activity_c, sleepdur_yest)) +
  geom_jitter()

ggplot(dat_test_sr, aes(hour_estimate, sleepdur_yest)) +
  geom_jitter() +
  geom_abline(intercept = 0, slope = 1) +
  labs(title = "Reported vs estimated sleep (jittered)", 
       x = "Hour estimate", y = "Self-reported sleep duration") +
  theme(text = element_text(size=16))

ggplot(dat_test_sr, aes(hour_estimate, SleepLNQuality)) +
  geom_jitter(alpha = 0.2) +
  labs(title = "Reported quality vs estimated duration (jittered)", 
       x = "Hour estimate", y = "Self-reported sleep quality") +
  theme(text = element_text(size=16))
```

Oura ring:

```{r fig.height=8}
n_presses_breaks = c(400, 1100, 3000, 8100, 22000)
ggplot(dat_test_oura, aes(hour_estimate, Oura_Sleep_hour)) +
  geom_jitter(aes(color = n_total_presses)) +
  geom_abline(intercept = 0, slope = 1) +
  scale_color_gradient(name = "Number of key presses", trans = "log",
                       breaks = n_presses_breaks, labels = n_presses_breaks) +
  labs(title = "Recorded vs estimated sleep (jittered)", 
       x = "Hour estimate", y = "Oura sleep duration") +
  theme(text = element_text(size=16))

ggplot(dat_test_oura, aes(hour_estimate, Oura_Sleep_hour)) +
  geom_jitter(aes(color = n_active_hours)) +
  geom_abline(intercept = 0, slope = 1) +
  labs(title = "Recorded vs estimated sleep (jittered)", 
       x = "Hour estimate", y = "Oura sleep duration") +
  theme(text = element_text(size=16))
```

Self-report and Oura ring:

```{r}
# 1887 observations
tmp <- dat_oura %>%
  drop_na(sleepdur_yest, Oura_Sleep_hour)

cor.test(tmp$sleepdur_yest, tmp$Oura_Sleep_hour, method = "spearman")
```

Overall:

```{r}
# Include all Oura and self-report data points, filter BiAffect data points
# based on train/test label
dat_descr <- dat %>%
  left_join(last_train_days, by = "subject") %>%
  select(subject, date, sleepdur_yest, Oura_Sleep_hour, hour_estimate, 
         label, dayNumber, last_train_day) %>%
  pivot_longer(sleepdur_yest:hour_estimate, names_to = "variable") %>%
  filter(variable != "hour_estimate" | (label == "test" & dayNumber > last_train_day + 3))

max_val = round(max(dat_descr$value, na.rm = TRUE))

g_violin <- ggplot(dat_descr, aes(variable, value, fill = variable)) +
  geom_violin() +
  geom_jitter(width = 0.05, height = 0, alpha = 0.1) +
  scale_x_discrete(labels = c("BiAffect", "Oura ring", "Self-report")) +
  scale_y_continuous(minor_breaks = seq(0, max_val)) +
  scale_fill_manual(values = c("lightcoral", "aquamarine3", "skyblue4")) +
  labs(x = NULL, y = "Sleep duration (hour)") +
  # facet_wrap(~ variable, scales = "free_y") +
  theme_minimal() +
  theme(
    text = element_text(size = 16),
    legend.position="none",
    axis.text.x = element_text(size = 16)
  )

g_violin

ggsave(file.path(man_img_dir, "violin_sleep.pdf"))

dat_descr %>%
  drop_na(value) %>%
  group_by(variable) %>%
  summarize(n())
```

```{r}
max_lim <- 19
r_coord = 15

tmp <- dat_descr %>%
  pivot_wider(id_cols = c(subject, date), 
              names_from = variable, values_from = value)

gs <- vector("list", length = 3)

gs[[1]] <- ggplot(tmp, aes(hour_estimate, Oura_Sleep_hour)) +
  labs(x = "BiAffect sleep duration (hour)", 
       y = "Oura sleep duration (hour)") +
  geom_label(aes(x, y), label = expression(r[s] == 0.40), parse = TRUE, 
             data = data.frame(x = r_coord, y = r_coord)) # Prevent overdrawing

gs[[2]] <- ggplot(tmp, aes(hour_estimate, sleepdur_yest)) +
  labs(x = "BiAffect sleep duration (hour)", 
       y = "Self-reported sleep duration (hour)") +
  geom_label(aes(x, y), label = expression(r[s] == 0.26), parse = TRUE, 
             data = data.frame(x = r_coord, y = r_coord)) # Prevent overdrawing

gs[[3]] <- ggplot(tmp, aes(sleepdur_yest, Oura_Sleep_hour))+
  labs(x = "Self-reported sleep duration (hour)", 
       y = "Oura sleep duration (hour)") +
  geom_label(aes(x, y), label = expression(r[s] == 0.74), parse = TRUE, 
             data = data.frame(x = r_coord, y = r_coord)) # Prevent overdrawing
  
gs <- lapply(gs, function(g) {
  g +
    geom_jitter(width = 0.2, height = 0.2, alpha = 0.5) +
    xlim(0, max_lim) +
    ylim(0, max_lim) +
    coord_fixed() +
    theme_minimal() +
    theme(text = element_text(size = 16))
})

g_scatter <- ggarrange(plotlist = gs, nrow = 1)

g_scatter
```

```{r fig.height=8}
ggarrange(g_violin, g_scatter, labels = c("A", "B"), font.label = list(size = 20), nrow = 2)

ggsave(file.path(man_img_dir, "violin_scatter.pdf"))
```

More than enough evidence for random slopes and intercepts:

```{r fig.height=8}
lmList <- nlme::lmList

m_list <- lmList(Oura_Sleep_hour ~ hour_estimate_c | subject, data = dat_test_oura, na.action = na.omit)

plot(intervals(m_list))
```

# Activity score

No correlation between random slope and intercept.

```{r}
m1 <- lmer(sleepdur_yest ~ activity_c + (activity_c || subject), dat_test_sr)

summary(m1)
```

There is structure in the residuals. This is because the response variable is discrete. Maybe ordinal regression makes more sense here.

```{r}
plot(m1)
```

Residuals also show some left-skew.

```{r}
qqnorm(resid(m1))
qqline(resid(m1))

hist(resid(m1))
```

Random effects look okay except for that outlier in the bottom left.

```{r}
re <- ranef(m1)$subject %>%
  rownames_to_column("subject") %>%
  pivot_longer(!subject)

ggplot(re, aes(sample = value)) +
  geom_qq() +
  geom_qq_line() +
  facet_wrap(~ name, scales = "free")
```

```{r fig.height=8}
plot(predict_response(m1, terms = "activity_c"), show_data = TRUE) +
  labs(title = "Predicted values of sleep duration", 
       x = "Activity (centred)", y = "Sleep duration") +
  theme(text = element_text(size = 16))
```

# Hour estimate

```{r}
cor.test(dat_test_sr$hour_estimate, dat_test_sr$sleepdur_yest, method = "spearman")
```

```{r}
m2_sr <- lmer(sleepdur_yest ~ hour_estimate_c + (hour_estimate_c | subject), 
              dat_test_sr)

summary(m2_sr)
```

To get the p values.

```{r paged.print=FALSE}
m2_nlme <- lme(sleepdur_yest ~ hour_estimate_c, 
               data = dat_test_sr, 
               random = ~ hour_estimate_c | subject)

summary(m2_nlme)
```


```{r}
plot(m2_sr)
```

Residuals also show some left-skew.

```{r}
qqnorm(resid(m2_sr))
qqline(resid(m2_sr))

hist(resid(m2_sr))
```

Random effects look okay, although the tails can get a little wonky.

```{r}
re <- ranef(m2_sr)$subject %>%
  rownames_to_column("subject") %>%
  pivot_longer(!subject)

ggplot(re, aes(sample = value)) +
  geom_qq() +
  geom_qq_line() +
  facet_wrap(~ name, scales = "free")
```

## Prediction

```{r fig.height=8}
plot(predict_response(m2_sr, terms = "hour_estimate_c"), show_data = TRUE, jitter = 0.1) +
  labs(title = "Predicted values of sleep duration", 
       x = "Hour estimate (centred)", y = "Sleep duration") +
  theme(text = element_text(size = 16))
```


```{r fig.height=8}
hour_mean <- dat_test_sr$hour_estimate_c %@% "scaled:center"
hour_sd <- dat_test_sr$hour_estimate_c %@% "scaled:scale"

pr <- predict_response(m2_sr, terms = "hour_estimate_c") %>%
  as.data.frame() %>%
  mutate(x = hour_sd * x + hour_mean)

n_presses_breaks = c(400, 1100, 3000, 8100, 22000)

max_est = round(max(dat_test_sr$hour_estimate))

g_pred_sr <- ggplot(pr, aes(x, predicted)) +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "grey") +
  geom_line(color = "lightcoral", lwd = 1.2) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), 
              fill="lightcoral", alpha = 0.2) +
  geom_jitter(
    aes(hour_estimate, sleepdur_yest, color = n_total_presses), 
    width = 0.2, height = 0.1, alpha = 0.5,
    data = dat_cmplt_sr) +
  scale_color_gradient(name = "Number of key presses", transform = "log",
                       breaks = n_presses_breaks, labels = n_presses_breaks, 
                       high = "aquamarine3") +
  scale_x_continuous(minor_breaks = seq(
    0, max_est
  )) +
  scale_y_continuous(minor_breaks = seq(0, 15), limits = c(0, 15)) + 
  labs(x = "BiAffect-estimated sleep duration (hour)", 
       y = "Self-reported sleep duration (hour)") +
  coord_fixed() +
  theme_minimal() +
  theme(
    text = element_text(size = 20),
    legend.title = element_text(size = 16),
    legend.text = element_text(size = 14)
  )

g_pred_sr

# ggsave(file.path(man_img_dir, "duration_prediction.pdf"))
```

```{r}
r_squared <- r2mlm(m2_sr)

# noquote class
vars <- r_squared$Decompositions[,1]
vars_df_sr <- data.frame(variance = unclass(vars)) %>%
  rownames_to_column("component") %>%
  filter(component != "fixed, between") %>%
  mutate(
    component = case_when(
      component == "fixed, within" ~ "Fixed effects",
      component == "slope variation" ~ "Random slope",
      component == "mean variation" ~ "Random intercept",
      component == "sigma2" ~ "Residuals"
    )
  )

vars_df_sr

cols <- rev(c("skyblue2", "skyblue3", "skyblue4", "grey25"))

g_var_sr <- ggplot(vars_df_sr, aes(1, variance)) +
  geom_col(aes(fill = fct_rev(component))) +
  scale_fill_manual(values = cols, name = "Variance component") +
  labs(x = NULL, y = "Proportion of variance") +
  scale_x_continuous(breaks = NULL) +
  scale_y_continuous(position = "right") +
  theme_minimal() +
  theme(
    text = element_text(size = 20),
    legend.title = element_text(size = 16),
    legend.text = element_text(size = 14)
  )

g_var_sr
```

```{r fig.height=8}
g_legends <- ggarrange(get_legend(g_pred_sr), get_legend(g_var_sr), ncol = 1, align = "v") 
g_plots <- ggarrange(g_pred_sr, g_var_sr, widths = c(7, 1), legend = "none")

ggarrange(g_plots, g_legends, widths = c(4, 1))

# ggsave(file.path(man_img_dir, "duration_prediction_sr_var.pdf"))
```

# Predicting Oura ring

```{r}
cortest <- cor.test(dat_test_oura$hour_estimate, dat_test_oura$Oura_Sleep_hour,
                    method = "spearman")

cortest
```

```{r}
tmp <- dat_test_oura %>%
  mutate(
    sedative_YN = as.integer(sedative_YN),
    stim_YN = as.integer(stim_YN),
    SSRI = as.integer(SSRI),
    SNRI = as.integer(SNRI),
    age_c = drop(age_c)
  )

m2_oura <- lmer(
  # Oura_Sleep_hour ~ hour_estimate_c + age_c + sedative_YN + stim_YN + SSRI + SNRI + (hour_estimate_c | subject),
  Oura_Sleep_hour ~ hour_estimate_c + (hour_estimate_c | subject),
  tmp)

summary(m2_oura)
```

```{r paged.print=FALSE}
m2_oura_nlme <- lme(Oura_Sleep_hour ~ hour_estimate_c, 
               data = dat_test_oura, 
               random = ~ hour_estimate_c | subject)

summary(m2_oura_nlme)
```

```{r}
get_coef_table <- function(m) {
  summ <- summary(m)
  tab <- summ$tTable
  tab <- tab[2:nrow(tab),]
  
  betas <- tab[,1]
  ses <- tab[,2]
  p_vals <- tab[,5]
  
  data.frame(beta = betas, SE = ses, p = p_vals)
}

# get_coef_table(m2_oura_nlme)
```


Cluster-centred predictors. Drawback is that the intercept and cluster (participant) means are almost perfectly correlated.

```{r}
m2_oura_cluster <- lmer(
  Oura_Sleep_hour ~ hour_estimate_cluster_centered + hour_estimate_cluster_mean + (hour_estimate_cluster_centered | subject),
  dat_test_oura)

summary(m2_oura_cluster)
```


## Evaluation

```{r}
plot(m2_oura)
```

```{r}
qqnorm(resid(m2_oura))
qqline(resid(m2_oura))

hist(resid(m2_oura))
```

Random effects look okay except for that outlier in the bottom left.

```{r}
re <- ranef(m2_oura)$subject %>%
  rownames_to_column("subject") %>%
  pivot_longer(!subject)

ggplot(re, aes(sample = value)) +
  geom_qq() +
  geom_qq_line() +
  facet_wrap(~ name, scales = "free")
```

```{r}
r_squared <- r2mlm(m2_oura)

# noquote class
vars <- r_squared$Decompositions[,1]
vars_df <- data.frame(variance = unclass(vars)) %>%
  rownames_to_column("component") %>%
  filter(component != "fixed, between") %>%
  mutate(
    component = case_when(
      component == "fixed, within" | component == "fixed" ~ "Fixed effects",
      component == "slope variation" ~ "Random slope",
      component == "mean variation" ~ "Random intercept",
      component == "sigma2" ~ "Residuals"
    )
  )

vars_df

cols <- rev(c("skyblue2", "skyblue3", "skyblue4", "grey20"))

g_var <- ggplot(vars_df, aes(1, variance)) +
  geom_col(aes(fill = fct_rev(component))) +
  scale_fill_manual(values = cols, name = "Variance component") +
  labs(x = NULL, y = "Proportion of variance") +
  scale_x_continuous(breaks = NULL) +
  scale_y_continuous(position = "right") +
  theme_minimal() +
  theme(
    text = element_text(size = 20),
    legend.title = element_text(size = 16),
    legend.text = element_text(size = 14)
  )

g_var
```

```{r}
r2mlm(m2_oura_cluster)
```


## Prediction

```{r fig.height=8}
plot(predict_response(m2_oura, terms = "hour_estimate_c"), show_data = TRUE, jitter = 0.1) +
  labs(title = "Predicted values of sleep duration", 
       x = "Hour estimate (centred)", y = "Sleep duration") +
  theme(text = element_text(size = 16))
```

```{r fig.height=8}
hour_mean <- dat_test_oura$hour_estimate_c %@% "scaled:center"
hour_sd <- dat_test_oura$hour_estimate_c %@% "scaled:scale"

pr <- predict_response(m2_oura, terms = "hour_estimate_c") %>%
  as.data.frame() %>%
  mutate(x = hour_sd * x + hour_mean)

n_presses_breaks = c(400, 1100, 3000, 8100, 22000)

max_est = round(max(dat_test_oura$hour_estimate))

g_pred <- ggplot(pr, aes(x, predicted)) +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "grey") +
  geom_line(color = "lightcoral", lwd = 1.2) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), 
              fill="lightcoral", alpha = 0.2) +
  geom_jitter(
    aes(hour_estimate, Oura_Sleep_hour, color = n_total_presses), 
    width = 0.2, height = 0.1,# alpha = 0.5,
    data = dat_test_oura) +
  scale_color_gradient(name = "Number of key presses", transform = "log",
                       breaks = n_presses_breaks, labels = n_presses_breaks, 
                       high = "aquamarine3") +
  scale_x_continuous(minor_breaks = seq(0, max_est)) +
  scale_y_continuous(minor_breaks = seq(0, 15), limits = c(0, 15)) + 
  labs(x = "BiAffect-estimated sleep duration (hour)", 
       y = "Oura-estimated sleep duration (hour)") +
  coord_fixed() +
  theme_minimal() +
  theme(
    text = element_text(size = 20),
    legend.title = element_text(size = 16),
    legend.text = element_text(size = 14)
  )

g_pred

# ggsave(file.path(man_img_dir, "duration_prediction_oura.pdf"))
```

```{r fig.height=8}
g_legends <- ggarrange(get_legend(g_pred), get_legend(g_var), ncol = 1, align = "v") 
g_plots <- ggarrange(g_pred, g_var, widths = c(7, 1), legend = "none")

ggarrange(g_plots, g_legends, widths = c(4, 1))

# ggsave(file.path(man_img_dir, "duration_prediction_oura_var.pdf"))
```

# Demographics

```{r}
ids <- as.integer(unique(dat_test_sr$subject))

dem_test <- dem %>%
  filter(id %in% ids) %>%
  select(!c(id, race_clear123, ethnicity, bmi, filing_status, sex_orient)) %>%
  select(where(~ !is.numeric(.x) || sum(.x, na.rm = TRUE) > 0)) %>%
  mutate(across(starts_with("currentdx"), as.logical))

tab_txt <- capture.output(CreateTableOne(data = dem_test)) %>%
  str_replace("^\\s\\s", "") %>%
  str_replace("(\\w|\\))\\s+(\\d.*)", "\\1;\\2") %>%
  str_replace_all("currentdx_(.+) = TRUE \\(%\\);(.*)", "   \\1;\\2")

cat(tab_txt, sep = "\n")
```

```{r}
ids <- as.integer(unique(dat_test_oura$subject))

dem_test <- dem %>%
  filter(id %in% ids) %>%
  select(!c(id, race_clear123, ethnicity, bmi, filing_status, sex_orient)) %>%
  select(where(~ !is.numeric(.x) || sum(.x, na.rm = TRUE) > 0)) %>%
  mutate(across(starts_with("currentdx"), as.logical))

tab_txt <- capture.output(CreateTableOne(data = dem_test)) %>%
  str_replace("^\\s\\s", "") %>%
  str_replace("(\\w|\\))\\s+(\\d.*)", "\\1;\\2") %>%
  str_replace_all("currentdx_(.+) = TRUE \\(%\\);(.*)", "   \\1;\\2")

cat(tab_txt, sep = "\n")
```

# Druijff-Van de Woestijne

```{r}
dat_dvdw <- read_parquet(file.path(dat_dir, "dat_dvdw.parquet"))

colnames(dat_dvdw)
```

```{r}
dat_dvdw_oura <- dat_dvdw %>%
  left_join(dat_oura %>% mutate(id = as.character(id)), 
            by = c("subject" = "id", "date" = "daterated")) %>%
  drop_na(Oura_Sleep_TST) %>%
  mutate(
    kap_c = drop(scale(kap)),
    n_active_hours_c = drop(scale(n_active_hours)),
    prev_n_active_hours_c = drop(scale(prev_n_active_hours))
  )

colnames(dat_dvdw_oura)
```

```{r}
m_dvdw <- lmer(Oura_Sleep_hour ~ kap_c + n_active_hours_c + prev_n_active_hours_c + (1 | subject),
               dat_dvdw_oura)

summary(m_dvdw)
```

```{r paged.print=FALSE}
m_dvdw_nlme <- lme(Oura_Sleep_hour ~ kap_c + n_active_hours_c + prev_n_active_hours_c, 
                   data = dat_dvdw_oura,
                   random = ~ 1 | subject,
                   na.action = na.omit)

summary(m_dvdw_nlme)
```


```{r}
plot(m_dvdw)

qqnorm(resid(m_dvdw))
qqline(resid(m_dvdw))

hist(resid(m_dvdw))

re <- ranef(m_dvdw)$subject %>%
  rownames_to_column("subject") %>%
  pivot_longer(!subject)

ggplot(re, aes(sample = value)) +
  geom_qq() +
  geom_qq_line() +
  facet_wrap(~ name, scales = "free")
```

```{r}
plot(predict_response(m_dvdw, terms = "kap_c"), show_data = TRUE) +
  xlim(-4, 4)
```


```{r fig.height=8}
kap_mean <- mean(dat_dvdw_oura$kap)
kap_sd <- sd(dat_dvdw_oura$kap)

pr <- predict_response(m_dvdw, terms = "kap_c") %>%
  as.data.frame() %>%
  mutate(x = kap_sd * x + kap_mean)

n_presses_breaks = c(400, 1100, 3000, 8100, 22000)

g_dvdw <- ggplot(pr, aes(x, predicted)) +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "grey") +
  geom_line(color = "lightcoral", lwd = 1.2) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), fill="lightcoral", alpha = 0.2) +
  geom_point(
    aes(kap, Oura_Sleep_hour),
    data = dat_dvdw_oura) +
  # scale_color_gradient(name = "Number of key presses", transform = "log",
  #                      breaks = n_presses_breaks, labels = n_presses_breaks, high = "aquamarine3") +
  scale_x_continuous(minor_breaks = ~ seq(0, 15)) +
  labs(x = "Keystroke-absence period (hour)", y = "Oura-estimated sleep duration (hour)") +
  theme_minimal() +
  theme(
    text = element_text(size = 20)
  )

# ggsave(file.path(man_img_dir, "duration_prediction_dvdw.pdf"))

g_dvdw
```

```{r}
r_squared <- r2mlm(m_dvdw)

# noquote class
vars <- r_squared$Decompositions[,1]
vars_df <- data.frame(variance = unclass(vars)) %>%
  rownames_to_column("component") %>%
  filter(component != "fixed, between", component != "slope variation") %>%
  mutate(
    component = case_when(
      component == "fixed" ~ "Fixed effects",
      component == "mean variation" ~ "Random intercept",
      component == "sigma2" ~ "Residuals"
    )
  )

vars_df

cols <- rev(c("aquamarine2", "aquamarine3", "grey20"))

g_var_dvdw <- ggplot(vars_df, aes(1, variance)) +
  geom_col(aes(fill = fct_rev(component))) +
  scale_fill_manual(values = cols, name = "Variance component") +
  labs(x = NULL, y = "Proportion of variance") +
  scale_x_continuous(breaks = NULL) +
  scale_y_continuous(position = "right") +
  theme_minimal() +
  theme(
    text = element_text(size = 20),
    legend.title = element_text(size = 16),
    legend.text = element_text(size = 14)
  )

g_var_dvdw
```

```{r fig.height=8}
ggarrange(g_dvdw, g_var_dvdw, widths = c(5, 2))

# ggsave(file.path(man_img_dir, "duration_prediction_oura_var_dvdw.pdf"))
```

